# Epigenomic data

## Introduction to `tidyCoverage`

### `CoverageExperiment` and `AggregatedCoverage` class

*Recommended time: 5 min*

Both classes are built on top of `SummarizedExperiment` class of objects. 

```{r eval = FALSE}
library(tidyCoverage)

data(ce)

data(ac)

ce

ac
```

### Creating a `CoverageExperiment` object from tracks and features

*Recommended time: 10 min*

A `CoverageExperiment` object extracts coverage scores from `tracks` (typically 
a named `BigWigFileList` or a list of already imported `RleList`s), over a list of 
genomic ranges of interest (typically a named `GRangesList`). 

```{r eval = FALSE}
tracks <- BigWigFileList(c(
    mnase = system.file("extdata", "MNase.bw", package = "tidyCoverage"), 
    cohesin = system.file("extdata", "Scc1.bw", package = "tidyCoverage")  
))
features <- GRangesList(
    TSSs = system.file("extdata", "TSSs.bed", package = "tidyCoverage") |> import() |> sample(2000), 
    TTSs = system.file("extdata", "TTSs.bed", package = "Bioc2024tidyWorkshop") |> import() |> sample(2000)
)

ce2 <- CoverageExperiment(
    tracks = tracks, 
    features = features,
    width = 2000, 
    ignore.strand = FALSE
)

ce2
```

Remember, this class of object is nearly identical to a `SummarizedExperiment` 
object. 

The slight difference is that each cell of the `coverage` assay is a `list` with 
a single element which is an array.  

```{r eval = FALSE}
colData(ce2)

rowData(ce2)

assay(ce2, 'coverage')

class(assay(ce2, 'coverage')['TSSs', 'mnase'])

class(assay(ce2, 'coverage')[['TSSs', 'mnase']])

dim(assay(ce2, 'coverage')[['TSSs', 'mnase']])
```

### Tidy coverage? That's right!

*Recommended time: 5 min*

Methods for `tidyverse` action verbs are available for `CoverageExperiment` objects, 
through the `tidySummarizedExperiment` package. 

```{r eval = FALSE}
library(tidySummarizedExperiment)

ce2

ce2 |> filter(features == 'TSSs')

ce2 |> slice(2)

ce2 |> select(features, n)
```

### `expand()` or `aggregate()`

*Recommended time: 5 min*

The `expand` method provided in the `tidyCoverage` package is used to 
turn a `CoverageExperiment` object into a tidy `tibble`. 

```{r eval = FALSE}
tib <- expand(ce2)

tib
```

The `aggregate` method provided in the `tidyCoverage` package is used to 
turn a `CoverageExperiment` object into an `AggregatedCoverage`. This class of 
object stores different statistical measurements of the coverage over 
each element of the `GRangesList` provided in the `CoverageExperiment` object. 

```{r eval = FALSE}
ac2 <- aggregate(ce2)

ac2
```

### Visualizing aggregated coverage

*Recommended time: 10 min*

The development version of `tidyCoverage` now ships two `geom_*` with 
appropriate default mapping aesthetics: 

- `geom_coverage()` for `CoverageExperiment` objects.
- `geom_aggrcoverage()` for `AggregatedCoverage` objects.

```{r eval = FALSE}
CoverageExperiment(tracks, GRanges("II:1-100000"), window = 100) |>
    expand() |>
    ggplot() +
    geom_coverage() + 
    facet_grid(track ~ features, scales = "free") + 
    labs(x = 'chrV', y = 'Signal coverage')

ggplot(ac2) + 
    geom_aggrcoverage() + 
    facet_grid(track ~ features, scales = "free") + 
    labs(x = 'Distance from genomic features', y = 'Signal coverage')
```

## Real-world use case: studying epigenomic landscape of reulatory elements

### Fetch coverage data from ENCODE

```{r eval = FALSE}
library(AnnotationHub)
ah <- AnnotationHub()
ids <- c('AH32207', 'AH35187')
names(ids) <- c('DNAse', 'H3K4me3')
bws <- lapply(ids, function(.x) ah[[.x]] |> resource()) |> BigWigFileList()
names(bws) <- names(ids)
```

### Plotting coverage data over several loci

```{r eval = FALSE}
ce3 <- CoverageExperiment(
    bws, 
    list(
        ccno = GRanges("chr5:55220001-55235000"),
        mcidas = GRanges("chr5:55159001-55174000")
    ), 
    window = 50
)
expand(ce3) |> 
    mutate(coverage = scales::oob_squish(coverage, c(0, 10))) |> 
    ggplot() +
    geom_coverage(aes(fill = track), unit = 'Mb') + 
    facet_grid(track~features, scales = 'free')
```

### Import DNase peaks from ENCODE

```{r eval = FALSE}
features <- list(DNase = ah[['AH30077']] |> filter(zScore > 100) |> sample(1000))
```

### Generating coverage aggregates and heatmaps over DNAse peak

```{r eval = FALSE}
ce4 <- CoverageExperiment(bws, features, width = 2000, window = 10)
```

```{r eval = FALSE}
aggregate(ce4) |> 
    ggplot(aes(x = coord, y = mean)) + 
    geom_aggrcoverage(aes(col = track)) + 
    facet_wrap(~track) + 
    labs(x = 'Distance from DNAse peak', y = 'Signal') 
```

```{r eval = FALSE, echo = FALSE}
ranges.order <- ce4 |> 
    expand() |> 
    filter(track == 'DNAse', coord.scaled >= -100, coord.scaled <= 100) |> 
    summarize(sum = sum(coverage, na.rm = TRUE)) |> 
    arrange(desc(sum)) |>
    pull(ranges) |> 
    rev()

ce4 |> 
    expand() |> 
    mutate(ranges = factor(ranges, levels = ranges.order)) |> 
    ggplot(aes(x = coord.scaled, y = ranges, fill = coverage)) + 
    geom_tile() +
    labs(x = 'Distance from DNAse peak', y = 'Signal') + 
    facet_grid(~track) +
    scale_x_continuous(expand = c(0, 0)) + 
    scale_fill_distiller(palette = "OrRd", na.value = '#fff5ed', direction = 1, limits = c(5, 60), oob = scales::squish) +
    theme_bw() + 
    theme(
        legend.position = 'top', 
        axis.ticks.y = element_blank(), axis.text.y = element_blank()
    )
``` 

# Resources {-}

# Session info {-}

::: {.callout-note collapse="true"}

## Click to expand 👇

```{r "session info"}
#| cache: false
sessioninfo::session_info(
    installed.packages()[,"Package"], 
    include_base = TRUE
)
```

:::

# References {-}
