# Manipulating genomic interval data

##  Importing `GRanges` from files

*Recommended time: 10 min*

The `BiocIO::import()` generic function lets one import `bed` files (or alike) 
into `GRanges` object in R. 

```{r eval = FALSE}
library(GenomicRanges)

library(rtracklayer)

bedf <- system.file('extdata', 'S288C-borders.bed', package = 'Bioc2024tidyWorkshop', mustWork = TRUE)

import(bedf)
```

The tidy way of importing genomic ranges in R, however, is to read files as 
a `data.frame` (or `tibble`) and _coerce_ it as a `GRanges` once it's imported in 
memory. 

```{r eval = FALSE}
library(tidyverse)

tib <- read_tsv(bedf, col_names = FALSE)

tib 

library(plyranges)

gr <- as_granges(tib, seqnames = X1, start = X2, end = X3)

gr
```

Note how refering to column names is done using [tidy evaluation](https://dplyr.tidyverse.org/articles/programming.html#data-masking). 

##  Manipulating `GRanges` with tidy verbs

*Recommended time: 10 min*

With `plyranges`, [a number of tidy operations](https://genomebiology.biomedcentral.com/articles/10.1186/s13059-018-1597-8/tables/1) are readily available, e.g.: 

- `mutate()`
- `select()`
- `filter()`
- `group_by()` + `summarize()`
- ... 

Just like with `tidyverse` functions, these operations work well with the native 
`|>` pipe in R. 

```{r eval = FALSE}
gr |> 
    mutate(score = runif(n())) |> 
    filter(score > 0.2) |> 
    mutate(round_score = round(score, digits = 1)) |> 
    group_by(round_score) |> 
    summarize(mean = mean(score))
```

But really, what `plyranges` excels at is providing direct access to "protected" 
variables, i.e. `seqnames`, `start`, `end`, `width`, `strand`, ...

```{r eval = FALSE}
gr |> 
    mutate(
        seqnames = factor('XVI', levels(seqnames)), 
        width = 1, 
        strand = rep(c('-', '+'), n()/2)
    )
```

Finally, a number of operations for genomic arithmetics are provided by `plyranges`: 

```{r eval = FALSE}
gr |> 
    anchor_center() |> 
    stretch(extend = -1000) |> 
    shift_upstream(250) |> 
    flank_upstream(100)
```

